---
title: "Simulation"
author: "Quinn White"
date: "`r Sys.Date()`"
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "./output")})
output: 
  rmdformats::readthedown:
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,message=FALSE, warning=FALSE)

library(tidyverse)
theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
  #  font <- "Arial"
    theme_bw() %+replace%    #replace elements we want to change
    
    theme(
      
      #text elements
      plot.title = element_text(             #title
                 #  family = font,            #set font family
                   size = 11,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),               
      
      plot.subtitle = element_text(          #subtitle
                #   family = font,            #font family
                   size = 10,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size
      
      axis.title = element_text(             #axis titles
                #   family = font,            #font family
                   size = 10),               #font size
      
      axis.text = element_text(              #axis text
                #   family = font,            #axis famuly
                   size = 8),
      strip.text = element_text(color="white", size = 9,
                                margin=unit(c(.1,.1,.1,.1), 'cm')),
      strip.background = element_rect(fill = "#363636"),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm")
      ) %+replace%
      theme(...)
   
}


```

Let $L_k$ be an indicator for having high blood pressure. Let $A$ be the treatment assigment (e.g., an ACE inhibitor).

Direct effect:

\begin{align*}
\text{Direct effect}&=\Pr[Y_{k+1}^{a=1, \overline{d=0}}=1]- \Pr[Y_{k+1}^{a=0, \overline{d=0}}=1] \\& = 0.5(0.2+0.1) - 0.5(.3+0.25)\\
&= -0.125
\end{align*}

Total effect:

\begin{align*}
\text{Total effect} &= \Pr[Y_{k+1}^{a=1} =1] - \Pr[Y_{k+1}^{a=0} = 1]
\end{align*}


```{r}

n <- 1000

set.seed(999)


baseline <- tibble(
  # baseline covariates
  l0 = rbinom(n, 1, .5),
  # treatment assigment
  a0 = rbinom(n, 1, .5),
  yk = 0,
  dk = 0,
  time_point = 0,
  C = 0,
  participant_id = 1:n)


# A1_L1_P_D <- 0
# A1_L0_P_D <- 0
# A0_L1_P_D <- 0
# A0_L0_P_D <- 0
# 
# 
# A1_L1_P_Y <- .8
# A1_L0_P_Y <- .8
# A0_L1_P_Y <- .2
# A0_L0_P_Y <- .2

```

```{r}

A1_L1_P_D <- .12
A1_L0_P_D <- .05
A0_L1_P_D <- .1
A0_L0_P_D <- .08


A1_L1_P_Y <- .15
A1_L0_P_Y <- .1
A0_L1_P_Y <- .2
A0_L0_P_Y <- .19




##########################
# FIRST TIME POINT 
########################## 


sim_dat <- map_df(1:n, ~ {
   baseline[.x,] %>%
     mutate(dk = case_when(
      a0 == 1 & l0 == 1 ~ rbinom(1,1, A1_L1_P_D),
      a0 == 1 & l0 == 0 ~ rbinom(1,1, A1_L0_P_D),
      a0 == 0 & l0 == 1 ~ rbinom(1,1,  A0_L1_P_D),
      a0 == 0 & l0 ==0 ~ rbinom(1, 1, A0_L0_P_D)
    ),
    dk_A_0 = case_when(
      # consistency 
      a0 == 0 ~ dk,
      # if a0 == 1 and l0 == 1, simulate as if a0 == 0 and l0 == 1
      a0 == 1 & l0 == 1 ~ rbinom(1,1, A0_L1_P_D),
      # if a0 == 1 and l0 == 0, simulate as if a0 == 0 and l0 == 0
      a0 == 1 & l0 == 0 ~ rbinom(1, 1, A0_L0_P_D)),
    dk_A_1 = case_when(
      # consistency
      a0 == 1 ~ dk,
      # if a0 == 0 and l0 == 1, simulate as if a0 == 1 and l0 ==1
      a0 == 0 & l0 == 1 ~ rbinom(1,1, A1_L1_P_D),
      # if a0 == 0 and l0 == 0, simulate as if a0 == 1 and l0 ==0
      a0 == 0 & l0 ==0 ~ rbinom(1,1, A1_L0_P_D)
    ),
    yk = case_when(
      # if they have the competing event, cannot have the event of interest
      dk == 1 ~ 0, 
      a0 == 1 & l0 == 1 ~ rbinom(1,1, A1_L1_P_Y),
      a0 == 1 & l0 == 0 ~ rbinom(1,1, A1_L0_P_Y),
      a0 ==0 & l0 == 1 ~ rbinom(1,1, A0_L1_P_Y),
      a0 == 0 & l0 ==0 ~ rbinom(1, 1, A0_L0_P_Y)
    ),
    yk_A_0 = case_when(
       # if they have the competing event, cannot have the event of interest
      # this remains true when considering a0 == 0
      dk == 1 ~ 0, 
      # consistency
      a0 == 0 ~ yk,
      # simulate as if they had a0 == 0 and l0 == 1
      a0 == 1 & l0 == 1 ~ rbinom(1,1,A0_L1_P_Y),
      # simulate as if they had a0 == 0 and l0 == 0
      a0 == 1 & l0 == 0 ~ rbinom(1,1, A0_L0_P_Y)
    ),
      yk_A_1 = case_when(
       # if they have the competing event, cannot have the event of interest
      # this remains true when considering a0 == 0
      dk == 1 ~ 0, 
      # consistency
      a0 == 1 ~ yk,
      # simulate as if they had a0 == 1 and l0 == 1
      a0 == 0 & l0 == 1 ~ rbinom(1,1,A1_L1_P_Y),
      # simulate as if they had a0 == 0 and l0 == 0
      a0 == 0 & l0 == 0 ~ rbinom(1,1, A1_L0_P_Y)
    ),
    yk_A_0_D_0 = case_when(
      # simulate as if they had no competing event and a0==0
      # consistency 
      dk == 0 & a0 == 0 ~ yk,
      # use previously generated potential outcomes when dk=0
      dk == 0 & a0 == 1 & l0 == 1 ~ yk_A_0,
      dk == 0 & a0 == 1 & l0 == 0 ~ yk_A_0,
      # when dk = 1 generate potential outcomes for these cases 
      dk == 1 & a0 %in% c(0,1) & l0 == 1 ~ rbinom(1,1, A0_L1_P_Y),
      dk == 1 & a0 %in% c(0,1) & l0 == 0 ~ rbinom(1,1, A0_L0_P_Y)
    ),
    yk_A_1_D_0 = case_when(
      # simulate as if they had no competing event and a0==1
      # consistency 
      dk == 0 & a0 == 1 ~ yk,
      # use previously generated potential outcomes when dk=0
      dk == 0 & a0 == 0 & l0 == 1 ~ yk_A_1,
      dk == 0 & a0 == 0 & l0 == 0 ~ yk_A_1,
      # when dk = 1 generate potential outcomes for these cases 
      dk == 1 & a0 %in% c(0,1) & l0 == 1 ~ rbinom(1,1, A1_L1_P_Y),
      dk == 1 & a0 %in% c(0,1) & l0 == 0 ~ rbinom(1,1, A1_L0_P_Y)
    )
    )})


##################################
# EFFECT OF L_0 ON L_1
##################################
L1_GIVEN_L0_P = .2
L1_GIVEN_L1_P = .6


###############################
# CENSORING PROBABILITIES
###############################
# probability of censoring if L_1 = 0
C_K1_GIVEN_L1 = .05
# probability of censoring if L_0 = 0
C_K1_GIVEN_L0 = .03


#################################################
# PROBABILITIES OF HAVING COMPETING EVENT 
# UNDER EACH TREATMENT x L1 x L0 COMBINATION 
#################################################
# first L is L_0, next is L_1
# e.g. L1_L0 indicates L_0=1, L_1=0
A1_L1_L0_P_D_K1 <- A1_L1_P_D + .04
A1_L0_L0_P_D_K1 <- A1_L0_P_D + .02
A1_L0_L1_P_D_K1 <- A0_L1_P_D + .06
A1_L1_L1_P_D_K1 <- A0_L0_P_D + .03

A0_L1_L0_P_D_K1 <- A1_L1_P_D + .02
A0_L0_L0_P_D_K1 <- A1_L0_P_D + .01
A0_L0_L1_P_D_K1 <- A0_L1_P_D + .03
A0_L1_L1_P_D_K1 <- A0_L0_P_D + .05

# A1_L1_L0_P_D_K1 <- A1_L1_P_D #+ .1
# A1_L0_L0_P_D_K1 <- A1_L0_P_D #+ .1
# A1_L0_L1_P_D_K1 <- A1_L1_P_D #+ .1 
# A1_L1_L1_P_D_K1 <- A1_L0_P_D #+ .1
# 
# A0_L1_L0_P_D_K1 <- A0_L1_P_D #+ .1
# A0_L0_L0_P_D_K1 <- A0_L0_P_D #+ .1
# A0_L0_L1_P_D_K1 <- A0_L1_P_D #+ .1
# A0_L1_L1_P_D_K1 <- A0_L0_P_D #+ .1


#########################################################
# PROBABILITIES OF HAVING EVENT OF INTEREST 
# UNDER EACH TREATMENT x L1 x L0 COMBINATION x L0 
#########################################################
# first L is L_0, next is L_1
# e.g. L1_L0 indicates L_0=1, L_1=0
A1_L1_L0_P_Y_K1 <- A1_L1_P_Y + .04
A1_L0_L0_P_Y_K1 <- A1_L0_P_Y + .03
A1_L0_L1_P_Y_K1 <- A1_L0_P_Y + 0.06
A1_L1_L1_P_Y_K1 <- A1_L1_P_Y + .08

A0_L1_L0_P_Y_K1 <- A0_L1_P_Y + .12
A0_L0_L0_P_Y_K1 <- A0_L0_P_Y +.1
A0_L0_L1_P_Y_K1 <- A0_L0_P_Y + .14
A0_L1_L1_P_Y_K1 <- A0_L1_P_Y + .16


################################
# ADD TIME POINT Y_{K+1}
################################
 sim_dat_updated <- map_df(1:n, ~{
   
   sim_dat[.x,] %>%
     mutate(l1 = case_when(
       # set to previous value if had either event of interest or competing event
       #dk == 1 | yk == 1 ~ l0,
       l0 == 0 ~ rbinom(1,1, L1_GIVEN_L0_P),
       l0 == 1 ~  rbinom(1,1, L1_GIVEN_L1_P)
       ),
       
      c_k1 = case_when(
        # if we observed either event previously, they are not censored
        dk == 1 | yk == 1 ~ 0,
        l1 == 1 ~ rbinom(1,1, C_K1_GIVEN_L1),
        l1 == 0 ~ rbinom(1,1, C_K1_GIVEN_L0)
        ),
      ############################
      # OBSERVED COMPETING EVENT
      ############################
      d_k1 = case_when(
        yk == 1 ~ 0,
        dk == 1 ~ 1,
        c_k1 == 1 ~ NA,
        
        # each combination of treatment x l1 x l0 
        a0 == 1 & l0 == 1 & l1 == 0 ~ rbinom(1,1,  A1_L1_L0_P_D_K1),
        a0 == 1 & l0 == 0 & l1 == 0 ~ rbinom(1,1,  A1_L0_L0_P_D_K1),
        a0 == 1 & l0 == 0 & l1 == 1 ~ rbinom(1,1,  A1_L0_L1_P_D_K1),
        a0 == 1 & l0 == 1 & l1 == 1 ~ rbinom(1,1,  A1_L1_L1_P_D_K1),
        a0 == 0 & l0 == 1 & l1 == 0 ~ rbinom(1,1,  A0_L1_L0_P_D_K1),
        a0 == 0 & l0 == 0 & l1 == 0 ~ rbinom(1,1,  A0_L0_L0_P_D_K1),
        a0 == 0 & l0 == 0 & l1 == 1 ~ rbinom(1,1,  A0_L0_L1_P_D_K1),
        a0 == 0 & l0 == 1 & l1 == 1 ~ rbinom(1,1,  A0_L1_L1_P_D_K1),
      ),
   
      
    ##############################################
    # COUNTERFACTUALS FOR COMPETING EVENT
    ##############################################
    # useful if we want to calculate effect of treatment on competing event
    
    # counterfactual under yk = 0, dk = 0, c_{k+1}=0, A=0 (needed for direct effect )
    d_k1_Y0_D0_C0_A_0 = case_when(
      # consistency 
      yk ==0 & dk == 0 & a0 == 0 & c_k1 == 0 ~ d_k1,
      # otherwise, have to generate new d_k1 
      # 1 combination for each possible history  (l0 x l1)
      a0 %in% c(0,1) & l0 == 1 & l1 ==0 ~ rbinom(1,1,   A0_L1_L0_P_D_K1),
      a0 %in% c(0,1) & l0 == 0 & l1 == 1 ~ rbinom(1, 1, A0_L0_L1_P_D_K1),
      a0 %in% c(0,1) & l0 == 0 & l1 == 0 ~ rbinom(1, 1, A0_L0_L1_P_D_K1),
      a0 %in% c(0,1) & l0 == 1 & l1 == 1 ~ rbinom(1, 1, A0_L1_L1_P_D_K1),
      ),
    
    # counterfactual under yk = 0, dk = 0, ck=0, A=1 (needed for direct effect ) 
    # eliminating possibility of competing event 
    d_k1_Y0_D0_C0_A_1 = case_when(
      # consistency 
      yk ==0 & dk == 0 & a0 == 1 & c_k1 == 0 ~ d_k1,
      # otherwise, have to generate new d_k1 
      # 1 combination for each possible history  (l0 x l1)
      a0 %in% c(0,1) & l0 == 1 & l1 == 1 ~ rbinom(1, 1, A1_L1_L1_P_D_K1),
      a0 %in% c(0,1) & l0 == 1 & l1 == 0 ~ rbinom(1, 1, A1_L1_L0_P_D_K1),
      a0 %in% c(0,1) & l0 == 0 & l1 == 1 ~ rbinom(1, 1, A1_L0_L1_P_D_K1),
      a0 %in% c(0,1) & l0 == 0 & l1 == 0 ~ rbinom(1, 1, A1_L0_L0_P_D_K1)
      ),
    
    # counterfactual under yk = 0, ck=0, a=0 (needed for total effect )
    d_k1_Y0_C0_A_0 = case_when(
      # set competing event to 1 if they had a previous event
      dk == 1 ~ 1,
      # consistency 
      yk == 0 & a0 == 0 & c_k1 == 0 ~ d_k1,
      # otherwise, use new the new outcome generated previously 
      TRUE ~ d_k1_Y0_D0_C0_A_0
      ),
    
    # counterfactual under yk=0, ck=0, a= 1 (needed for total effect )
    d_k1_Y0_C0_A_1 = case_when(
      # set competing event to 1 if they had a previous event
      dk == 1 ~ 1,
      # consistency 
      yk ==0 & a0 == 0 & c_k1 == 0 ~ d_k1,
      # otherwise, use new the new outcome generated previously 
      TRUE ~ d_k1_Y0_D0_C0_A_1
      ),
    
    
    ##################################
    # OBSERVED EVENT OF INTEREST
    ##################################
     y_k1 = case_when(
       #  yk == 1 ~ 1,
        # dk == 1 ~ 0,
        c_k1 == 1 ~ NA,
        # each combination of treatment x l1 x l0 
        a0 == 1 & l0 == 1 & l1 == 0 ~ rbinom(1,1, A1_L1_L0_P_Y_K1),
        a0 == 1 & l0 == 0 & l1 == 0 ~ rbinom(1,1, A1_L0_L0_P_Y_K1),
        a0 == 1 & l0 == 0 & l1 == 1 ~ rbinom(1,1, A1_L0_L1_P_Y_K1),
        a0 == 1 & l0 == 1 & l1 == 1 ~ rbinom(1,1, A1_L1_L1_P_Y_K1),
        a0 == 0 & l0 == 1 & l1 == 0 ~ rbinom(1,1, A0_L1_L0_P_Y_K1),
        a0 == 0 & l0 == 0 & l1 == 0 ~ rbinom(1,1, A0_L0_L0_P_Y_K1),
        a0 == 0 & l0 == 0 & l1 == 1 ~ rbinom(1,1, A0_L0_L1_P_Y_K1),
        a0 == 0 & l0 == 1 & l1 == 1 ~ rbinom(1,1, A0_L1_L1_P_Y_K1),
      ),
    # set censoring to zero
    y_k1_A_0 = case_when(
      yk == 1 ~ 1,
      dk == 1  | d_k1 ==1 ~ 0,
      # consistency
      a0 == 0 & c_k1 == 0 ~ y_k1,
      l0 == 1 & l1 == 0 ~ rbinom(1,1, A0_L1_L0_P_Y_K1),
      l0 == 0 & l1 == 0 ~ rbinom(1,1, A0_L0_L0_P_Y_K1),
      l0 == 0 & l1 == 1 ~ rbinom(1,1, A0_L0_L1_P_Y_K1),
      l0 == 1 & l1 == 1 ~ rbinom(1,1, A0_L1_L1_P_Y_K1)
    ),
    
     y_k1_A_1 = case_when(
      yk == 1 ~ 1,
      dk == 1 | d_k1 ==1 ~ 0,
      # consistency
      a0 == 1 & c_k1 == 0 ~ y_k1,
      l0 == 1 & l1 == 0 ~ rbinom(1,1, A1_L1_L0_P_Y_K1),
      l0 == 0 & l1 == 0 ~ rbinom(1,1, A1_L0_L0_P_Y_K1),
      l0 == 0 & l1 == 1 ~ rbinom(1,1, A1_L0_L1_P_Y_K1),
      l0 == 1 & l1 == 1 ~ rbinom(1,1, A1_L1_L1_P_Y_K1)
    ),
    
    ##############################################
    # COUNTERFACTUALS FOR EVENT OF INTEREST 
    ##############################################

    # counterfactual under yk = 0, d_{k+1} = 0, c_{k+1}=0, A=0 (needed for direct effect)
    y_k1_Y0_D0_C0_A_0 = case_when(
      # consistency 
      yk ==0 & d_k1 == 0 & a0 == 0 & c_k1 == 0 ~ y_k1,
      # otherwise, have to generate new y_k1 
      # 1 combination for each possible history  (l0 x l1)
      a0 %in% c(0,1) & l0 == 1 & l1 == 1 ~ rbinom(1, 1, A0_L1_L1_P_Y_K1),
      a0 %in% c(0,1) & l0 == 1 & l1 == 0 ~ rbinom(1, 1, A0_L1_L0_P_Y_K1),
      a0 %in% c(0,1) & l0 == 0 & l1 == 1 ~ rbinom(1, 1, A0_L0_L1_P_Y_K1),
      a0 %in% c(0,1) & l0 == 0 & l1 == 0 ~ rbinom(1, 1, A0_L0_L0_P_Y_K1)
      ),
    
    # counterfactual under y_{k} = 0, d_{k+1} = 0, c_{k+1}=0, A=1 (needed for direct effect)
    # eliminates possibility of competing event
    y_k1_Y0_D0_C0_A_1 = case_when(
      # consistency 
      yk ==0 & d_k1 == 0 & a0 == 1 & c_k1 == 0 ~ y_k1,
       # otherwise, have to generate new y_k1 
      # 1 combination for each possible history  (l0 x l1)
      a0 %in% c(0,1) & l0 == 1 & l1 == 1 ~ rbinom(1, 1, A1_L1_L1_P_Y_K1),
      a0 %in% c(0,1) & l0 == 1 & l1 == 0 ~ rbinom(1, 1, A1_L1_L0_P_Y_K1),
      a0 %in% c(0,1) & l0 == 0 & l1 == 1 ~ rbinom(1, 1, A1_L0_L1_P_Y_K1),
      a0 %in% c(0,1) & l0 == 0 & l1 == 0 ~ rbinom(1, 1, A1_L0_L0_P_Y_K1)
      ),
    
    # counterfactual under y_{k} = 0, c_{k+1}=0, A=0 (needed for total effect)
    y_k1_Y0_C0_A_0 = case_when(
      # we are not eliminating possbility of past competing event
      dk == 1 ~ 0, 
      # consistency 
      yk == 0 & a0 == 0 & c_k1 == 0 ~ y_k1,
     # otherwise, use new outcome generated
      TRUE ~ y_k1_Y0_D0_C0_A_0
      ),

    # counterfactual under y_{k} = 0, c_{k+1}=0, A=1 (needed for total effect)
    y_k1_Y0_C0_A_1 = case_when(
      # we are not eliminating possibility of past competing event
      dk == 1 ~ 0, 
      # consistency 
      yk ==0 & a0 == 1 & c_k1 == 0 ~ y_k1,
      TRUE ~ y_k1_Y0_D0_C0_A_1
      )
     )
 })


# eliminate rows for anyone who already had an event at time k
sim_dat_updated <- sim_dat_updated %>% filter(! (yk ==1 | dk ==1 ))
 
# mean is proportion
mean(sim_dat_updated$y_k1_A_1, na.rm=TRUE) - mean(sim_dat_updated$y_k1_A_0, na.rm=TRUE)

mean(sim_dat_updated$y_k1_A_1, na.rm=TRUE)
mean(sim_dat_updated$y_k1_A_0, na.rm=TRUE)

```


```{r}

########################################################################
# REFORMATTING FOR COMPATIBILITY WITH gfoRmula package
########################################################################

sim_dat_k_formatted <- sim_dat %>% 
  select(l0, a0, dk, yk, participant_id, time_point) %>%
  mutate(C=0) %>%
  bind_rows(baseline) %>%
  rename(L = l0,
         A = a0,
         D =dk,
         Y = yk )

 sim_dat_k1_formatted <- sim_dat_updated %>% 
   select(c_k1, d_k1, y_k1, participant_id, a0) %>%
   # L_2 is not used in the estimation 
   mutate(time_point = 2,
          L = NA) %>%
   rename(A=a0,
          D=  d_k1,
          Y = y_k1,
          C= c_k1) 
 
 sim_dat_formatted <- sim_dat_k_formatted %>% 
   rbind(sim_dat_k1_formatted) %>%
   arrange(participant_id, time_point) %>%
   as.data.frame()
 
 
```
 
 
 
```{r}
 
histvars <- c('L')
histories <- c('lagged')
covparams <- list(covmodels = list(c(L ~ A + lag1_L)))


basecovs <- c('A')
  
 
outcome_model <- Y ~ A + L + lag1_L
compevent_model <- D ~ A + L + lag1_L
intvars <- c('A', 'A')
interventions <- list( list(static, rep(0, 2)),
                       list(static, rep(0, 2))) 
 

 ```
 
 

```{r}
 
 library(gfoRmula)
 
 
 # we obtain total effect when competing event is not treated as a censoring event
 # (set compevent_cens = FALSE)
 ###################
 # TOTAL EFFECT
  ###################
 result <- gformula(obs_data = sim_dat_formatted,
                    id = 'participant_id',
                    time_name = 'time_point',
                    covnames = c('L'),
                    covtypes=c('binary'),
                    covparams = list(covmodels = list(c(L ~ A + lag1_L))),
                    histvars = list(c('L')),
                    histories = c(lagged),
                    # treatment is not time-varying
                    basecovs = 'A',
                    outcome_name = 'Y',
                    outcome_type = 'survival',
                    compevent_name = 'D',
                    ymodel =  c(Y ~ A + L + lag1_L),
                    compevent_model = c(D ~ A + L + lag1_L),
                    # total effect
                    compevent_cens = FALSE,
                    censor_name = 'C',
                    censor_model = c(C ~ L),
                    intvars = list('A', 'A'),
                    interventions = list( list(c(static, rep(0, 3))),
                                         list(c(static, rep(1, 3)))))
 
 
# sim_dat_formatted_dt <- data.table::data.table(sim_dat_formatted)
 
 
 
sim_dat_formatted_dt <- sim_dat_formatted %>% 
  group_by(participant_id) %>%
  mutate(Y = lead(Y, n=1),
         D = lead(D, n=1),
         C = lead(C, n=1)) %>% 
  filter(time_point != 2) %>% 
  ungroup() %>%
  as.data.frame() %>%
  mutate(Y = ifelse(D == 1, NA, Y)) %>%
  data.table::data.table()
 
 sim_dat_formatted_dt <- sim_dat_formatted_dt %>%
   mutate(Y = ifelse(is.na(Y), 0, Y))
 
 
 sim_dat_formatted_dt %>%
   count(A, Y, D,L, time_point, C) |> arrange(Y)

 
 sim_dat_formatted_dt %>% cor(use="complete.obs")

result <- gformula(obs_data = sim_dat_formatted_dt,
                    id = 'participant_id',
                    time_name = 'time_point',
                    time_points = 2,
                    covnames = c('L'),
                    basecovs = c('A'),
                    covtypes=c('binary'),
                    covparams = list(covlink = c('logit'),
                                     covmodels = c(L ~ A + lag1_L)),
                    histvars = list(c('L')),
                    histories = c(lagged),
                    # treatment is not time-varying
                    # basecovs = 'A',
                    outcome_name = 'Y',
                    outcome_type = 'survival',
                    compevent_name = 'D',
                    ymodel =  Y ~ A + L + lag1_L,
                    compevent_model = D ~ A + L + lag1_L,
                    # total effect
                    compevent_cens = FALSE,
                    ref_int=1,
                   int_descript=c('Never Treat', 'Always Treat'),
                    censor_name = 'C',
                    censor_model = C ~ L,
                    intvars = list(c('A'), c('A')),
                    interventions = list( list(c(static, rep(0, 2))),
                                          list(c(static, rep(1, 2)))),
                    seed =999
                    )
  

  
  
 
 
 
 
 # we obtain direct effect when competing event is not treated as a censoring event
 # (set compevent_cens = TRUE)
 ###################
 # TOTAL EFFECT
  ###################
 result <- gformula(obs_data = sim_dat_formatted,
                    id = 'participant_id',
                    time_name = 'time_point',
                    covnames = 'L',
                    covtypes='binary',
                    covparams =  list(covlink = c('logit'),
                                      covmodels = list(c(L ~ A + lag1_L))),
                    histvars = 'L',
                    histories = 'lagged',
                    # treatment is not time-varying
                    basecovs = 'A',
                    outcome_name = 'Y',
                    outcome_type = 'survival',
                    compevent_name = 'D',
                    ymodel = outcome_model,
                    compevent_model = D ~ A + L + lag1_L,
                    # total effect
                    compevent_cens = FALSE,
                    intvars = c('A', 'A'),
                    interventiosn = list( list(static, rep(0, 2)),
                       list(static, rep(0, 2))))
 
 

```


```{r, eval=FALSE}
        


label_a <- function(string, prefix = "A=") paste0(prefix, string)
label_l <- function(string, prefix = "L=") paste0(prefix, string)


# sim_dat %>%
#   pivot_longer(cols=contains("k")) %>%
#   group_by(name,l0, a0) %>%
#   summarize(prop = sum(value) /n()) %>%
#   ggplot(aes(y= name, x = prop)) +
#   geom_bar(stat="identity") +
#   facet_grid(l0~a0, labeller = labeller(a0=label_a, l0 = label_l)) +
#   theme_c()



sim_dat %>%
  pivot_longer(cols=contains("k")) %>%
  group_by(name,l0, a0) %>%
  mutate(a0 = ifelse(a0 == 0, "Control", "Treatment")) %>%
  summarize(prop = sum(value) /n()) %>%
  ggplot(aes(y= name, x = prop, fill = factor(a0))) +
  geom_bar(stat="identity", position="dodge", alpha=.8) +
  facet_wrap(~l0, labeller = labeller( l0 = label_l)) +
  theme_c() +
  labs(fill="", x= "Proportion who had Event") +
  viridis::scale_fill_viridis(discrete=TRUE, 
                              begin=.1, 
                              end=.8) 



```


True average causal effect

```{r}
# for k = 0, l0=1

### A = 0
## K = 0
# L0 = 1
EY_A0 <-
  A0_L1_P_Y * (1-A0_L1_P_D) * 0.5 +
# L0 = 0
  A0_L0_P_Y * (1-A0_L0_P_D) * 0.5 +
## K = 1
# L0 = 0, L1 = 0
  A0_L0_L0_P_Y_K1 * (1-A0_L0_L0_P_D_K1) * (1-L1_GIVEN_L0_P) * (1-A0_L0_P_Y) * (1-A0_L0_P_D) * 0.5 +
# L0 = 1, L1 = 1
  A0_L1_L1_P_Y_K1 * (1-A0_L1_L1_P_D_K1) * (L1_GIVEN_L1_P) * (1-A0_L1_P_Y) * (1-A0_L1_P_D) * 0.5 +
# L0 = 0, L1 = 1
  A0_L0_L1_P_Y_K1 * (1-A0_L0_L1_P_D_K1) * (L1_GIVEN_L0_P) * (1-A0_L0_P_Y) * (1-A0_L0_P_D) * 0.5 +
# L0 = 1, L1 = 0
  A0_L1_L0_P_Y_K1 * (1-A0_L1_L0_P_D_K1) * (1-L1_GIVEN_L1_P) * (1-A0_L1_P_Y) * (1-A0_L1_P_D) * 0.5 


### A = 1
## K = 0
# L0 = 1
EY_A1 <-
  A1_L1_P_Y * (1-A1_L1_P_D) * 0.5 +
# L0 = 0
  A1_L0_P_Y * (1-A1_L0_P_D) * 0.5 +
## K = 1
# L0 = 0, L1 = 0
  A1_L0_L0_P_Y_K1 * (1-A1_L0_L0_P_D_K1) * (1-L1_GIVEN_L0_P) * (1-A1_L0_P_Y) * (1-A1_L0_P_D) * 0.5 +
# L0 = 1, L1 = 1
  A1_L1_L1_P_Y_K1 * (1-A1_L1_L1_P_D_K1) * (L1_GIVEN_L1_P) * (1-A1_L1_P_Y) * (1-A1_L1_P_D) * 0.5 +
# L0 = 0, L1 = 1
  A1_L0_L1_P_Y_K1 * (1-A1_L0_L1_P_D_K1) * (L1_GIVEN_L0_P) * (1-A1_L0_P_Y) * (1-A1_L0_P_D) * 0.5 +
# L0 = 1, L1 = 0
  A1_L1_L0_P_Y_K1 * (1-A1_L1_L0_P_D_K1) * (1-L1_GIVEN_L1_P) * (1-A1_L1_P_Y) * (1-A1_L1_P_D) * 0.5 

EY_A1
EY_A0
EY_A1 - EY_A0

EY_A1
EY_A0


```

```{r}

# ### A = 1
# ## K = 0
# # L= 1
# A1_L1_P_Y * (1-A1_L1_P_D) * .5
# # L = 0
# A1_L0_P_Y * (1-A1_L0_P_D) * .5
# 
# 
# 0.2 * 1* (1-0.12) * 0.5 
# # for k = 0, l0=0
# 0.1 * 1 *(1-.1) * .5 

# for k=1
```


* Find package to estimate total and direct effect 
* Restructure data 
* Compare IPW and G-compare 
* Make changes to simulation and compare 

R package gformula 

